In this worksheet we consider planar statics problems, and how to solve the
linear systems of equations that arise when solving them.

From Newton's laws of motion, for the static equilibrium of a body both
translational forces and moments must balance. Specifically, the vector sum of
the external forces must equal zero:
\begin{equation} \label{eq:trans}
 \sum_{i} {\bf F}_{i}= \bf{0}
\end{equation}
This is called the {\em external force balance}. Also, the sum of moments,
$\bf{M}$, about an arbitrary point ${\bf r}$ of all forces is zero:
\begin{equation} \label{eq:mom}
 \sum_{i} {\bf M}_{i}= \sum_{i} ({\bf r}_{i}-{\bf r}) \times {\bf F}_{i}= 0.
\end{equation}
This is called the {\em external torque balance}. It can be proven that there
is no additional linearly independent balance solely involving the external
forces.

Now for planar (2D) statics problems,~\eqref{eq:trans} becomes
\[
 {\bf F}_{i}=\begin{pmatrix}
         F_{i,x} \\ F_{i,y} \\ 0
        \end{pmatrix}
\]
so it follows that the external force balance yields two equations, $\sum
F_{i,x} =0$ and $\sum F_{i,y}=0$.

For such problems the moment of any force will be in the $z$ direction so it
only has one non-zero component $M_3$. Also we can find the $z$-moment $M_3$
by simply multiplying the force by the perpendicular distance through which it
levers. The external torque balance~\eqref{eq:mom} yields a single equation,
so there are three equations in total for static equilibrium.

For a given statics problem with more unknowns than equations, the system of
simultaneous equations does not have a unique solution and the problem is said
to be {\em statically indeterminate}.

If one body is {\em resting} on another, the normal reaction force $N$ must be
positive.

If friction acts a point of contact, the friction condition $F\leq \mu N$ must
always be satisfied.

For the purpose of calculating moments we can always consider that the force of
gravity acting on a body acts at its \emph{centre of mass}. The centre of mass
can be defined formally but for symmetric bodies we can simply think of this
as the ``middle'' of the body.

%The force in a spring of stiffness $k$ with extension $x$ from its natural
%length $L$ is given by Hooke's law, $F=kx$.

\clearpage
